---
title: "A1"
author: "Kerui Du"
date: "`r lubridate::today()`"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

### 1a
```{r 1a}
library(s20x)
PPW.df <- read.csv('data/PPW.csv')
plot(Wt~Ht, type="n", data=PPW.df)     
text(Wt~Ht, labels=substr(How,1,1),
     col=ifelse(How=="Drown","blue", "brown"),
     data=PPW.df, cex=0.9, font=2)

```

By looking at the scatter plot, there is a positive linear relationship between the person's height and the weight of the person's lungs. As height increases, the lungs' weight increases as well, which means taller people tend to have heavier lungs. In the meantime, trend lines for the manner of death***(drown or hung)*** are both positive and seems to parallel**(we can test this by fitting an interaction term later)**; in that case, it means that the height effect appears the same regardless of the manner of death. 

However, it is evident that people who are the same height are more likely to die from drowning than hanging. No wonder that people who died from drowning will have larger lungs. The reason is that it has water on them.


### 1b
```{r 1b}
plot(Wt~Ht, type="n", data=PPW.df) 
text(Wt~Ht, labels=substr(How,1,1),
     col=ifelse(How=="Drown","blue", "brown"),
     data=PPW.df, cex=0.9, font=2)
text(x = 196, y = 1125, labels = 'p', cex = 1.5, col = 'red')

```

Fitting a point with height(195cm) and lungs weight(1125g). It is pretty hard to distinguish which group this point belongs to at first glance, as it lies in the middle of two different manners of death.

This point relates to prosecution and defense counsels' is that as a prosecution role, they will emphasize that point P died from drowning even though it is relatively more minor than the same height. On the contrary, defense counsel will insist that point p died by hanging since this point is consistent with the trend line of hung.


### 1c
```{r 1c}
inter <- lm(Wt~Ht*How, data = PPW.df)
plot(inter, which = 1)
anova(inter)
summary(inter)

no_inter <- lm(Wt~Ht+How, data = PPW.df)
plot(no_inter, which = 1)
anova(no_inter)
normcheck(no_inter)
cooks20x(no_inter)
summary(no_inter)
```
Method and Assumption checks:
To begin with a linear model fitted the model with explanatory variables height, how, and their interaction. However, the interaction term was not significant(p-value = 0.2809). Therefore, the model was refitted with the interaction term removed. 

For assumption check, there was a curvature in the residual plot, which means the quadratic term can be implemented, whereas most of the data were reasonably constant, which was still able to stick to linear. Most of the data points were along with the line in QQ plot and histogram was normally distributed; also, there were no points of undue influence. Hence, all model assumptions were satisfied.

Our final model is 
                  Wti = β0 + β1 × Hti + β2 × HowHungi + εi,
                  where HowHungi set to one if person i died by hanging, otherwise it's zero,
                  and εi∼iid N(0,σ2).



### 1d
```{r 1d}
n <- nrow(PPW.df)
new.df <- PPW.df[c(1:n, n+1), ]
new.df[n + 1, c('Ht', 'Wt')] = c(196, 1125)

no_inter_new <- lm(Wt~Ht+How, data = new.df)
c <- coef(no_inter_new)
plot(Wt~Ht, type="n", data=PPW.df) 
text(Wt~Ht, labels=substr(How,1,1),
     col=ifelse(How=="Drown","blue", "brown"),
     data=PPW.df, cex=0.9, font=2)
text(x = 196, y = 1125, labels = 'p', cex = 1.5, col = 'red')
abline(c[1], c[2], col = 'blue', lty = 2)
abline(c[1] + c[3], c[2], col = 'blue', lty = 2)
```

It is clear that two lines are going up and parallel.

As a role of the prosecution, point P likely is one of the people who died from drowning. The reason is that point P is a data that belongs to drown but with relatively large variability.

As a role of defence counsel, they may have strong evidence against prosecution as point P is perfectly matched on the line(for hung). Hence, they could declare that point P died by hanging.



### 1e
```{r 1e}
preds.df <- data.frame(How = c('Drown', 'Hung'), Ht = 196)
(preds <- predict(no_inter_new, preds.df, interval = 'prediction'))
```



### 1f
```{r 1f}
plot(Wt~Ht, type="n", data=PPW.df) 
text(Wt~Ht, labels=substr(How,1,1),
     col=ifelse(How=="Drown","blue", "brown"),
     data=PPW.df, cex=0.9, font=2)
text(x = 196, y = 1125, labels = 'p', cex = 1.5, col = 'red')
abline(c[1], c[2], col = 'blue', lty = 2)
abline(c[1] + c[3], c[2], col = 'blue', lty = 2)
segments(196-.5,preds[1,2], 196-.5, preds[1,3], col="green")
segments(196+.5,preds[2,2], 196+.5, preds[2,3], col="blue")
```

### 1g
For a person whose height is 196cm and weight 1125g died from drowning, we predict the weight of this person's lungs is between 980.0 to 1670.0(gram), as a role of the prosecution, it is an acceptable value as this point is somewhere in the prediction interval***(green line. However)***, if this person died by hanging, we predict the weight of lungs is between 750.6 to 1443.6(gram), as a role of defence counsels, it is can be an evidence since this point is in the middle of the prediction interval***(blue line)***.



### 1h
```{r 1h}
glm_fit <- glm(Wt~Ht+How, data = new.df, family = 'gaussian')
RSS <- sum(resid(glm_fit)^2); RSS
R_squared <- (glm_fit$null.deviance - RSS) / glm_fit$null.deviance; R_squared

```



## Question 2
### 2a
```{r 2a}
PPW.df$yesno <- with(PPW.df, ifelse(How=='Hung', 1, 0))
bin <- glm(yesno~Wt+Ht, data = PPW.df, family = 'binomial'); bin
```

### 2b
```{r 2b}
pred.interval <- predict(bin, data.frame(Ht = 196, Wt = 1125), se.fit = T)
lower <- pred.interval$fit-1.96*pred.interval$se.fit
upper <- pred.interval$fit+1.96*pred.interval$se.fit
ci <- cbind(lower, upper)
round(exp(ci) / (1+exp(ci)), digits = 3)
```

### 2c
We predict that the probability of PPW with height is 196cm and weight 1125g, which died by hanging is between 32.5% to 72.5% on average.

As a role of the prosecution, the confident interval of probability for point P is partly greater than 50%. It is not plausible to prove that PPW died by hanging.

As a role of defence counsel, more than half of the confident interval of probability is over 50%. It is reasonable to trust that PPW died by hanging.


### 2d
Since the confident interval of the probability that PPW died by hanging is between 32.5% to 72.5%, that is, it is somehow strong evidence against that PPW died from drowning, meanwhile, we can back to the plot we were refitted the point P and obverse that P is probably died by hanging, in this case, we possibly assume an accused is innocent.

