---
title: "mid-test"
author: "Kerui Du"
date: "16/09/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1
### (a)
model function:

log($u_{i}$) = $\beta_{0}$ + $\beta_{1}*ReserveHahei_{i}$ + $\beta_{2}*ReserveTAWH_{i}$ 
+ $\beta_{3}*SeasonSpring{i}$ + $\beta_{4}*Year98-99{i}$

$Y_{i} \sim {\sf Poisson}(u_{i})$

where Reserve at Hahei when the $i^{th}$ of $ReserveHahei_{i}$=1 and Reserve at TAWH when the $i^{th}$ of $ReserveTAWH_{i}$=1, Otherwise Reserve at CROP, Year is between 1998-1999 when the $i^{th}$ of $Year98-99{i}$= 1, otherwise is between 1997-1998, Season is spring when the $i^{th}$ $SeasonSpring{i}$=1, otherwise, Season is autumn. $Y_{i}$ is the number of snapper in the recording collected by the $i^{th}$ camera.


### (b)
```{r 1b}
exp(c(-.928897, -.6669))
100*(exp(c(-.928897, -.6669))-1)
```

We estimate that cameras deployed in spring detect between 0.39 and 0.51 times as many snapper
as those deployed in autumn.

In the CROP Reserve and Year between 1997 to 1998, with 95% confidence interval, we estimate that the season spring is somewhere between 48.7% to 60.5% lower than the season autumn 



## Q1 

### (a)
model function:

logit($u_{i}$) = $\beta_{0}$ + $\beta_{1}*wantsMoreyes_{i}$ + $\beta_{2}*age25-29_{i}$ 
+ $\beta_{3}*age30-39_{i}$ + $\beta_{4}*age40-49_{i}$ + $\beta_{5}*educationlow_{i}$

$Y_{i} \sim {\sf Binomial}(n_{i}, p_{i})$

where carving for more children is Yes when the $i^{th}$ of $wantsMoreyes_{i}$=1, otherwise is 0, 
the education level is low when the $i^{th}$ of $educationlow_{i}$=1, the age of woman is 25 to 29 when the $i^{th}$ of $age25-29_{i}$=1, the age of woman is 30 to 39 when the $i^{th}$ of $age30-39_{i}$=1, or the age of woman is 40 to 49 when the $i^{th}$ of $age40-49_{i}$=1, otherwise is 0, $n_{i}$ is total number of woman who using contraception for $i^{th}$ group, and $Y_{i}$ is the 

## Q2 

### (a) 
The null hypothesis being tested by the anova is that the linear model is appropriate, we don't any 
interaction term between variables age and education.

###(b)
We don't have evidence to suggest our null hypothesis is not appropriate, since the p-value is greater than 0.05, therefore, we may stick with bin1.fit(without interaction). 


## Q3

### (a)
No, we have strong evidence to suggest the model is not appropriate. If the model is correct, we
might expect the deviance to come from a chi-squared distribution with 10 degrees of freedom,
but this is not plausible.


### (b)
No, we have strong evidence to suggest the model is not appropriate. If the model is correct, we
might expect the deviance to come from a chi-squared distribution with 7 degrees of freedom,
but this is not plausible.

```{r}
p = 0.16231/(1+0.16231)
var1 = 264*p*(1-p);var1

2.8288*var1

var1*(1+264*0.0085823-0.0085823)

(42.849-52*p)/sqrt(var1)

(38.348-52*p)/sqrt(var1*(1+264*0.0085823-0.0085823))
```






